---
title: "elec_consumption"
output: pdf_document
date: "2026-08-03"
---
We have the electricity consumption in kW that we want to forecast. 
We have also the outdoor temperature of the building. 
the data is collected every 15min from 1/1/2010 1:15 until 2/18/2010 23:45

```{r}
library(readxl)
df = read_excel("Elec-train.xlsx")
#rename column 
names(df)[names(df) == "Power (kW)"] = "power"
names(df)[names(df) == "Temp (C°)"] = "temp"
names(df)[names(df) == "Timestamp"] = "time"

df$time = as.POSIXct(df$time, format = "%m/%d/%Y %H:%M")
# we work with the data from 1/1/2010 1:15 until 2/18/2010 23:45
df = df[df$time <= as.POSIXct("2/18/2010 23:45", format = "%m/%d/%Y %H:%M"),]

head(df)
```

```{r}
library(xts)
elect = xts(df$power, order.by = df$time)
plot(
  elect,
  xaxt = "n",
  type = "l",
  main = "15-minute Electricity Consumption",
  xlab = "Day",
  ylab = "Electricity"
)

# add daily labels
days <- seq(
  as.POSIXct("2010-01-01"),
  as.POSIXct("2010-02-18"),
  by = "1 day"
)

axis(
  1,
  at = days,
  labels = format(days, "%b-%d"),
  las = 2
)
```

```{r}
library(dplyr)
library(ggplot2)


# Create date and time-of-day variables
df_seasonal <- df %>%
  mutate(
    date = as.Date(time),
    time_minutes = as.numeric(format(time, "%H")) * 60 +
                   as.numeric(format(time, "%M"))
  )

# Seasonal plot
ggplot(df_seasonal,
       aes(x = time_minutes,
           y = power,
           group = date,
           colour = date)) +
  geom_line(linewidth = 0.5, alpha = 0.7) +
  scale_x_continuous(
    breaks = seq(0, 1440, by = 120),
    labels = sprintf("%02d:%02d",
                     seq(0, 1440, by = 120) %/% 60,
                     seq(0, 1440, by = 120) %% 60)
  ) +
  labs(
    title = "Daily Seasonal Pattern of Electricity Consumption",
    x = "Time of Day",
    y = "Electricity",
    colour = "Date"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```
We saw that there is daily pattern in this dataset but there are some drops at the beginning of 18th Feb. so i decided to take the last two days for testing so that we don't see a big error for the drop
```{r}
filter(df,power==0)
```

#Spliting dataset
```{r}
# split the last two days for testing models 
df_train = df[df$time <= as.POSIXct("2/16/2010 23:45", format = "%m/%d/%Y %H:%M"),]
df_test = df[df$time > as.POSIXct("2/16/2010 23:45", format = "%m/%d/%Y %H:%M"),]
x_train = ts(df_train$power, start = c(1,5),frequency = 24*4)
```




